---
output:
  html_document:
    theme: readable
    df_print: paged
    highlight: tango
---

```{r setup, include=FALSE}
# set global doc options:
# do not include warning messages
# in the knitted document
knitr::opts_chunk$set( echo=T, message=F, warning=F, image.width=10)
```
**Introducing montyhall**

**Abstract**

The Monty Hall game (probability problem) follows a game show format where 
the player is given the option to select ONE of THREE doors. Similar to other 
hidden doors (hidden prizes) games, the player does not know what is behind 
each door. Two doors have goats behind them. One door has a new car behind it. 
The player gets the option of choosing a door, but the prize is not yet revealed. 
The host then opens one of the two doors—always revealing a goat. Lastly, 
the player has one final choice, stay with their first choice or choose 
the last remaining door.

create_game() generates a new game that consists of two doors with goats 
behind them, and one with a car.

**Description**

The game setup replicates the game on the TV show "Let's
Make a Deal" where there are three doors for a contestant
to choose from, one of which has a car behind it and two 
have goats. The contestant selects a door, then the host
opens a door to reveal a goat, and then the contestant is
given an opportunity to stay with their original selection
or switch to the other unopened door. There was a famous 
debate about whether it was optimal to stay or switch when
given the option to switch, so this simulation was created
to test both strategies. 

The function returns a length 3 character vector
indicating the positions of goats and the car.

```{r}

```

**Create_game**
```{r}

# step 1: create a vector of 3 doors: 2 goats and 1 car
# step 2: randomize the position of the car for a new game
# step 3: return the new game vector 

create_game <- function( )
{
    prizes <- c("goat","goat","car")
  a.game <- sample( x=prizes, size=3, replace=F )
    return( a.game )
}

```

**Next step**

The contestant chooses one door, but no prizes are revealed.

"DoorS" vector created for contestant--1, 2, 3.
The sample function is then used to randomly pick
one of three doors. This simulates the contestant's
choice.


The function returns the number of the 
randomly selected door.

**Select_door**
```{r}

#Step 5: create a vector of doors numbered 1,2,3
#Step 6: randomly select ONE of the doors
#Step 7: return the selection

#Changes made to my code based on solutions
#Removed "replace=F"
#Replaced "C(1,2,3)" woth doors in sample

select_door <- function( )
{

  doors <- c(1,2,3) 
  a.pick <- sample( x=doors, size=1 )
  return( a.pick )

}

```

**Next step**

The host opens one of the doors containing a goat. 
If the contestant has already selected a goat
in their first pick, the host will open the other
goat door. If the contestant has picked the door
with the car, the host will open one or two
goat doors revealing a goat.

If the contestant's first pick is a car,
the host will open one of two doors containing
a goat. If the contestant's first pick is a goat,
the host will open the only remaining door with
a goat.

The contestant will be given the number of a
door containing a goat. This is the door
opened by the host.

**Open_goat_door**
```{r}

#Step 8: The door can't be a door the contestant has already selected.

#Changes made based on Lab 1 solutions... add open_goat_door before other_door function
#If the host has two doors to choose from, the host will still need to pick ONE.
#Open_goat_door function should be based on game, a.pick instead of first.game, select.door
#Add doors <- c(1,2,3) and IF statements
#Remove other_door and other.door vectors

open_goat_door <- function( game, a.pick )

{
    doors <- c(1,2,3)
   # if contestant selected car,
   # randomly select one of two goats 
   if( game[ a.pick ] == "car" )
   { 
     goat.doors <- doors[ game != "car" ] 
     opened.door <- sample( goat.doors, size=1 )
   }
   if( game[ a.pick ] == "goat" )
   { 
     opened.door <- doors[ game != "car" & doors != a.pick ] 
   }
   return( opened.door ) # number between 1 and 3
  
}

```

**Next step**
The contestant either stays with their first choice,
Or they select the door not yet opened by the host.

Knowing that one goat door has been revealed, the
contestant doesn't yet know which door has a car.
They can either stay with their first choice or
choose the one remaining unpicked and unrevealed door.

The contestant will be given the number of a
door containing a goat. This is the door
opened by the host.

**Change_door**
```{r}

#Step 9: Change Doors?

#First option: Contestant decides to stay with first choice.
#Changes made based on Lab 1 solutions... update change_door functions to include IF statements

change_door <- function( stay=T, opened.door, a.pick )
{
  
doors <- c(1,2,3) 
   
   if( stay )
   {
     final.pick <- a.pick
   }
   if( ! stay )
   {
     final.pick <- doors[ doors != opened.door & doors != a.pick ] 
   }
  
   return( final.pick )  # number between 1 and 3
  

}

```

**Next step**

The prized is revealed based on the 
contestant's final door selection.

If the contestant's final pick contains
a goat, they lose the game.
If the contestant's final pick contains
a car, they win the game.

**Determine_winner**
```{r}

#Step 10: Test both options.

#Changes made based on Lab 1 solutions... update determine_winner functions to reflect change_door IF statements

determine_winner <- function( final.pick, game )
{
   if( game[ final.pick ] == "car" )
   {
      return( "WIN" )
   }
   if( game[ final.pick ] == "goat" )
   {
      return( "LOSE" )
   }
}

```

Play_game contains all of the aforementioned code and functions
and allows users to create a new game.

**Play_game**
```{r}

play_game <- function( )
{
  new.game <- create_game()
  first.pick <- select_door()
  opened.door <- open_goat_door( new.game, first.pick )
  final.pick.stay <- change_door( stay=T, opened.door, first.pick )
  final.pick.switch <- change_door( stay=F, opened.door, first.pick )
 outcome.stay <- determine_winner( final.pick.stay, new.game  )
  outcome.switch <- determine_winner( final.pick.switch, new.game )

  # game.results <- bundle the results
  # return( <<< game.results >>> )
  
  strategy <- c("stay","switch")
  outcome <- c(outcome.stay,outcome.switch)
  game.results <- data.frame( strategy, outcome,
                              stringsAsFactors=F )
  return( game.results )
}

```

```{r}

play_game()

```

Game loop. This for-loop allows users to play
multiple games and review results.

Users can review the results of multiple
games using a for-loop. This will allow the user
to compare wins to loses to help understand 
probability.

All functions combined to create a new game,
multiple times.

**For_loop**
```{r}

play_n_games <- function( n=100 )
{
  
  library( dplyr )
  results.list <- list()   # collector
  loop.count <- 1

  for( i in 1:n )  # iterator
  {
    game.outcome <- play_game()
    results.list[[ loop.count ]] <- game.outcome 
    loop.count <- loop.count + 1
  }
  
  results.df <- dplyr::bind_rows( results.list )

  table( results.df ) %>% 
  prop.table( margin=1 ) %>%  # row proportions
  round( 2 ) %>% 
  print()
  
  return( results.df )

}

```

```{r}

play_n_games()

```
